Precise predictions for same-sign W-bosons scattering at the LHC

### 2018-01-18 Draft/contributions.tex
Update WHIZARD LO cross section in the Draft
also fix typos in shower.tex and contributions.tex.

At tree level, there are three contributions to the $W^+W^+$ production in association with two jets: the pure EW component $\mathcal{O}(\alpha^6)$, the QCD background $\mathcal{O}(\alphas^2\alpha^4)$ and the interference $\mathcal{O}(\alphas\alpha^5)$.\\
The three contributions to the the jet--jet invariant mass $M_{jj}$ and positive rapidity difference $|\Delta y_{jj}|$ distributions are shown in Fig.~\ref{fig:mjjdyjj_1d}, together with their sum.
\begin{figure*}[hbt]
\centering
\includegraphics[scale=0.395]{figures/scanfigures/mjj_full.pdf}
\includegraphics[scale=0.395]{figures/scanfigures/dyjj_full.pdf}
\caption{Differential distribution in the jet--jet invariant mass $M_{jj}$ (left) and the positive difference of the jet rapidities $|\Delta y_{jj}|$ (right) at LO. EW contribution in red, QCD in green, interference in gray and sum in blue. No cuts on $M_{jj}$ and $|\Delta y_{jj}|$ are applied. Results of full LO matrix element simulation.}
\label{fig:mjjdyjj_1d}
\end{figure*}
Concerning the QCD contribution, it is worth noticing that for quantum numbers conservation external gluons cannot contribute at LO, thus the $\mathcal{O}(\alphas^2\alpha^4)$ diagrams only involve $t/u$--channel virtual gluons: this results in a small QCD background, although the VBF topological cuts are not imposed. The interference between EW and QCD contributions is small, due to color suppression, but not negligible ($t/u$ interference with identical fermions). The LO cross--section for the three perturbative orders are shown in table~\ref{tab:LOscanXsec}.
\begin{table}[h!]
    \centering
    \begin{tabular}{c|c|c|c}
        Code  &  $\sigma[\rm{fb}] (\mathcal{O}(\alpha^6))$ & $\sigma[\rm{fb}] (\mathcal{O}(\alphas^2\alpha^4))$ & $\sigma[\rm{fb}] (\mathcal{O}(\alphas\alpha^5))$ \\
        \hline
        \hline
        {\sc Phantom}  &  $ 2.292 \pm 0.002 $ & $ 1.477 \pm 0.001 $ & $ 0.223 \pm 0.003 $ \\
        {\sc Xxx}&  $ \pm $ & $ \pm $ & $ \pm $
    \end{tabular}
    \caption{\label{tab:LOscanXsec} Cross sections at LO accuracy for the ${\rm p}{\rm p}\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}{\rm j}{\rm j}$ process: these results concern the setup described in section~\ref{subsec:inputpar}, apart from the $M_{jj}$ and $|\Delta y_{jj}|$ cuts which have been dropped.}
\end{table}
The EW, QCD and interference amount respectively to 57\%, 37\% and 6\% of the total inclusive cross section. In the presence of typical VBS cuts, the EW contribution is strongly enhanced, as shown in section~\ref{subsec:LOfiducial}.\\
The EW contribution to the $M_{jj}$ distribution (red curve in Fig.~\ref{fig:mjjdyjj_1d}) shows a huge peak around the $W$ mass: this is due to the possibility for the two jets to be decay products of a $W^-$ boson. Note that this effect doesn't arise in calculations which employ VBS approximation (see VBS approx. description in section [ref.]).
One can also see this in three different plots in the plan $\left(m_{\Pj\Pj}, \Delta y_{\Pj\Pj}\right)$.
\begin{figure}[ht]
\centering
\includegraphics[scale=0.395]{figures/scanfigures/scan_ew6.pdf}
\includegraphics[scale=0.395]{figures/scanfigures/scan_ew5qcd1.pdf}
\includegraphics[scale=0.395]{figures/scanfigures/scan_ew4qcd2.pdf}
\caption{Cross sections (fb) per bin in the plan $\left(m_{\Pj\Pj}, \Delta y_{\Pj\Pj}\right)$ for the three LO contributions of orders $\mathcal{O}(\alpha^6)$ (top left), $\mathcal{O}(\alpha^5\alphas)$ (top right), and $\mathcal{O}(\alpha^4 \alphas^2)$ (bottom).}
\label{fig:mjjdyjj_2d_LO}
\end{figure}

>>>>

At tree level, there are three contributions to the $W^+W^+$ production in association with two jets: the pure EW component $\mathcal{O}(\alpha^6)$, the QCD background $\mathcal{O}(\alphas^2\alpha^4)$ and the interference $\mathcal{O}(\alphas\alpha^5)$.\\
The three contributions to the the jet--jet invariant mass $M_{jj}$ and positive rapidity difference $|\Delta y_{jj}|$ distributions are shown in Fig.~\ref{fig:mjjdyjj_1d}, together with their sum.
\begin{figure*}[hbt]
\centering
\includegraphics[scale=0.395]{figures/scanfigures/mjj_full.pdf}
\includegraphics[scale=0.395]{figures/scanfigures/dyjj_full.pdf}
\caption{Differential distribution in the jet--jet invariant mass $M_{jj}$ (left) and the positive difference of the jet rapidities $|\Delta y_{jj}|$ (right) at LO. EW contribution in red, QCD in green, interference in gray and sum in blue. No cuts on $M_{jj}$ and $|\Delta y_{jj}|$ are applied. Results of full LO matrix element simulation.}
\label{fig:mjjdyjj_1d}
\end{figure*}
Concerning the QCD contribution, it is worth noticing that for quantum number conservation external gluons cannot contribute at LO, thus the $\mathcal{O}(\alphas^2\alpha^4)$ diagrams only involve $t/u$--channel virtual gluons: this results in a small QCD background, although the VBF topological cuts are not imposed. The interference between EW and QCD contributions is small, due to color suppression, but not negligible ($t/u$ interference with identical fermions). The LO cross--section for the three perturbative orders are shown in table~\ref{tab:LOscanXsec}.
\begin{table}[h!]
    \centering
    \begin{tabular}{c|c|c|c}
        Code  &  $\sigma[\rm{fb}] (\mathcal{O}(\alpha^6))$ & $\sigma[\rm{fb}] (\mathcal{O}(\alphas^2\alpha^4))$ & $\sigma[\rm{fb}] (\mathcal{O}(\alphas\alpha^5))$ \\
        \hline
        \hline
        {\sc Phantom}  &  $ 2.292 \pm 0.002 $ & $ 1.477 \pm 0.001 $ & $ 0.223 \pm 0.003 $ \\
        {\sc Xxx}&  $ \pm $ & $ \pm $ & $ \pm $
    \end{tabular}
    \caption{\label{tab:LOscanXsec} Cross sections at LO accuracy for the ${\rm p}{\rm p}\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}{\rm j}{\rm j}$ process: these results concern the setup described in section~\ref{subsec:inputpar}, apart from the $M_{jj}$ and $|\Delta y_{jj}|$ cuts which have been dropped.}
\end{table}
The EW, QCD and interference amount respectively to 57\%, 37\% and 6\% of the total inclusive cross section. In the presence of typical VBS cuts, the EW contribution is strongly enhanced, as shown in section~\ref{subsec:LOfiducial}.\\
The EW contribution to the $M_{jj}$ distribution (red curve in Fig.~\ref{fig:mjjdyjj_1d}) shows a huge peak around the $W$ mass: this is due to the possibility for the two jets to be decay products of a $W^-$ boson. Note that this effect doesn't arise in calculations which employ VBS approximation (see VBS approx. description in section [ref.]).
One can also see this in three different plots in the plan $\left(m_{\Pj\Pj}, \Delta y_{\Pj\Pj}\right)$.
\begin{figure}[ht]
\centering
\includegraphics[scale=0.395]{figures/scanfigures/scan_ew6.pdf}
\includegraphics[scale=0.395]{figures/scanfigures/scan_ew5qcd1.pdf}
\includegraphics[scale=0.395]{figures/scanfigures/scan_ew4qcd2.pdf}
\caption{Cross sections (fb) per bin in the plan $\left(m_{\Pj\Pj}, \Delta y_{\Pj\Pj}\right)$ for the three LO contributions of orders $\mathcal{O}(\alpha^6)$ (top left), $\mathcal{O}(\alpha^5\alphas)$ (top right), and $\mathcal{O}(\alpha^4 \alphas^2)$ (bottom).}
\label{fig:mjjdyjj_2d_LO}
\end{figure}

### 2018-01-18 Draft/crosssectionLO.tex
some edits for the shower part
Update WHIZARD LO cross section in the Draft
also fix typos in shower.tex and contributions.tex.

 In Tab.~\ref{tab:wg1_LOrates} we report the total rates at LO accuracy obtained with the set-up described above, and in Fig.~\ref{fig:wg1_mjj-llLO} we show the results
for the tagging-jet (left) and lepton-pair (right) invariant-mass distribution. In both case we show the absolute distributions in the main frame of the 
figures, while in the inset the ratio over {\sc VBFNLO} is displayed. For both observables we find 
an excellent agreement among the various tools, which confirms the fact
that contributions from $s-$channel diagrams as well as from non-resonant configurations are strongly suppressed in the fiducial region.
\begin{table}[h!]
    \centering
    \begin{tabular}{c|c}
        Code  &  $\sigma[\rm{fb}]$  \\
        \hline
        \hline
        {\sc Bonsay}  &  $X \pm 0.0002$ \\
        {\sc MG5\_aMC}&  $X \pm 0.001$  \\ 
        {\sc MoCaNLO+Recola}  &  $1.4347 \pm 0.0001$ \\
        {\sc PHANTOM}&  $1.4374 \pm 0.0006 $  \\
        {\sc POWHEG}  &  $1.44092 \pm 0.00009$ \\
        {\sc VBFNLO}  &  $1.43796 \pm 0.00005$ \\
        {\sc Whizard}&  $1.4363 \pm 0.0009 $
    \end{tabular}
    \caption{\label{tab:wg1_LOrates} Rates at LO accuracy within VBS cuts obtained with the different codes used in this comparison, 
    for the ${\rm p}{\rm p}\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}{\rm j}{\rm j}$ process.}
\end{table}

>>>>

 In Tab.~\ref{tab:wg1_LOrates} we report the total rates at LO accuracy obtained with the set-up described above, and in Fig.~\ref{fig:wg1_mjj-llLO} we show the results
for the tagging-jet (left) and lepton-pair (right) invariant-mass distribution. In both cases we show the absolute distributions in the main frame of the 
figures, while in the inset the ratio over {\sc VBFNLO} is displayed. For both observables we find 
an excellent agreement among the various tools, which confirms the fact
that contributions from $s-$channel diagrams as well as from non-resonant configurations are strongly suppressed in the fiducial region. We have checked 
that the same level of agreement holds for many other differential distributions.
\begin{table}[h!]
    \centering
    \begin{tabular}{c|c}
        Code  &  $\sigma[\rm{fb}]$  \\
        \hline
        \hline
        {\sc Bonsay}  &  $X \pm 0.0002$ \\
        {\sc MG5\_aMC}&  $X \pm 0.001$  \\ 
        {\sc MoCaNLO+Recola}  &  $1.4347 \pm 0.0001$ \\
        {\sc PHANTOM}&  $1.4374 \pm 0.0006 $  \\
        {\sc POWHEG}  &  $1.44092 \pm 0.00009$ \\
        {\sc VBFNLO}  &  $1.43796 \pm 0.00005$ \\
        {\sc Whizard}&  $1.4360 \pm 0.0005 $
    \end{tabular}
    \caption{\label{tab:wg1_LOrates} Rates at LO accuracy within VBS cuts obtained with the different codes used in this comparison, 
    for the ${\rm p}{\rm p}\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}{\rm j}{\rm j}$ process.}
\end{table}

### 2018-01-18 Draft/shower.tex
Update WHIZARD LO cross section in the Draft
also fix typos in shower.tex and contributions.tex.

We now turn to discuss how different predictions compare when the matching to parton-shower (PS) is included. For such
a comparison we expect larger discrepancy than what we found at fixed-order, as a consequence of the different
matching schemes, PS employed and of the other details of the matching (such as the choice of the shower initial scale). Among
the codes capable of providing fixed-order results, presented before, {\sc MG5\_aMC}, {\sc POWHEG} and {\sc VBFNLO}
can also provide results at (N)LO+PS accuracy. Besides, also {\sc PHANTOM} is employed for LO+PS results.\\
{\sc MG5\_aMC},
which
employs the {\sc MC@NLO}~\cite{Frixione:2002ik} matching procedure, will be used together with {\sc Pythia8}~\cite{Sjostrand:2014zea} (version 2.2.3) 
and {\sc Herwig++}~\cite{Bahr:2008pv, Bellm:2013hwb} (version 2.7.1). For {\sc POWHEG}, the omonymous matching procedure is 
employed~\cite{Nason:2004rx,Frixione:2007vw}, together with {\sc Pythia8} 
{\bf MZ VERSION? if same as MG5, put it at the end together with the tune}. {\sc VBFNLO} makes it possible to choose between the {\sc MC@NLO} and {\sc POWHEG}
matching, in both cases together with {\sc Herwig7}~\cite{}. Finally, {\sc PHANTOM} results will be shown matched with {\sc Pythia 8}.
Whenever {\sc Pythia8}\ is used, the Monash tune~\cite{Skands:2014pea} is selected.\\

Results will be presented within the cuts described in Section~\ref{subsec:inputpar}, applied after shower and hadronization (this implies that jets
are obtained by clustering stable hadrons, and not QCD partons). This implies that at the event-generation level, looser cuts (or no cuts at all) 
must be emplouyed in order not to bias the results. {\bf MZ lepton-jet separation at the hard-event level?}.\\

>>>>

We now turn to discuss how different predictions compare when the matching to parton-shower (PS) is included. For such
a comparison we expect larger discrepancy than what we found at fixed-order, as a consequence of the different
matching schemes, PS employed and of the other details of the matching (such as the choice of the shower initial scale). Among
the codes capable of providing fixed-order results, presented before, {\sc MG5\_aMC}, {\sc POWHEG} and {\sc VBFNLO}
can also provide results at (N)LO+PS accuracy. Besides, also {\sc PHANTOM} is employed for LO+PS results.\\
{\sc MG5\_aMC},
which
employs the {\sc MC@NLO}~\cite{Frixione:2002ik} matching procedure, will be used together with {\sc Pythia8}~\cite{Sjostrand:2014zea} (version 2.2.3) 
and {\sc Herwig++}~\cite{Bahr:2008pv, Bellm:2013hwb} (version 2.7.1). For {\sc POWHEG}, the anonymous matching procedure is 
employed~\cite{Nason:2004rx,Frixione:2007vw}, together with {\sc Pythia8} 
{\bf MZ VERSION? if same as MG5, put it at the end together with the tune}. {\sc VBFNLO} makes it possible to choose between the {\sc MC@NLO} and {\sc POWHEG}
matching, in both cases together with {\sc Herwig7}~\cite{}. Finally, {\sc PHANTOM} results will be shown matched with {\sc Pythia 8}.
Whenever {\sc Pythia8}\ is used, the Monash tune~\cite{Skands:2014pea} is selected.\\

Results will be presented within the cuts described in Section~\ref{subsec:inputpar}, applied after shower and hadronization (this implies that jets
are obtained by clustering stable hadrons, and not QCD partons). This implies that at the event-generation level, looser cuts (or no cuts at all) 
must be employed in order not to bias the results. {\bf MZ lepton-jet separation at the hard-event level?}.\\
